An Approach to Off-line Handwriting Recognition through Segmentation Techniques and Deep Neural Networks

### 2017-12-17 Report/paper.tex
go

\subsection{Contouring}
Our text segmentation approach is based on finding contours in the text image.
We can find these contours using %TODO:: reference paper contouring suzuki85
These contours in postprocessing stages to recover the original words from the text image.
The contouring algorithm used is the one implemented in the opencv library for python.
This algorithm generates a lot of unneeded contours, which we will filter through postprocessing.
Also since words in handwritten text are not always connected into one piece, we will need to connect these contours in the postprocessing stages as well.

\subsection{Postprocessing}
In the postprocessing step, we eliminate the redundant extracted contours, and reconnect the contours of words that consist of multiple parts.

\subsubsection{Contour removal}
Since the contouring algorithm produces contours that are not contours of the words in the text, these contours need to be filtered out and removed.
To accomplish this we make use of the bounding box we can draw around a contour.
This bounding box is the smallest rectangle containing said contour, and it will allow us to easily compare contour positions in the text image.

The first filter is removing the contours of which the bounding box is contained by the bounding box of another contour.
This is allowed because the contents of this inner bounding box will be regarded when processing the contents of the outer bounding box, and so we can remove the contours belonging to the inner bounding boxes.

The second filter is removing the contours that are too small to contain parts of a word.
Since the contouring algorithm extracts the contours for any size, and because our character recognition network does not process punctuation,
we need to eliminate these from our results.

We achieve this by taking the average height of all the found contours in the text.
From these contours we calculate the average height, after which we remove the contours with a height lesser than the average height multiplied by a chosen parameter $minimalHeightMultiplier$.
The optimal results were achieved when we initialize with $minimalHeightMultiplier = 2/3$, so that we retain only contours higher than the average height $ * 2/3$.

\subsubsection{Contour combination}
With the excess contours removed, we can begin to reconstruct the original words from the text by combining multiple contours into one where needed.
The algorithm used for this is a straightforward overlapping algorithm, that combines word segments based on the overlap between the bounding box of their contours.
The parameters for this algorithm are the $alowedWordGap$ and the $allowedVerticalOverlap$ paramters.
These parameters are initialized as $interWordSegmentDistance = $ the average width of the found words $ / 16 $ and $allowedVerticalOverlap = 1/2$ to give the best results in the our testset, however further experimentation with these parameters and the algorithm is required to find a more generalized optimal solution.

@@

\subsection{Line selection}
Now that the words are extracted from the text image, the next goal is to try and reconstruct the original text by reordering these words in their original lines.
For the line selection process, an iterative algorithm, Algorithm \ref{alg:linereconstr}, is used.
\begin{algorithm}[tb]
   \caption{Algorithm for line reconstruction}
   \label{alg:linereconstr}
\begin{algorithmic}
   \STATE {\bfseries Input:} $extractedWords$, $textlines$
   \REPEAT
        \STATE $highestWord = findHighestWord $
        \STATE $newline = [highestWord]$
        \STATE Initialize $noChange = true$.
        \STATE $avgY= newline$.findAverageYpos
        \REPEAT
            \FOR{$word$ {\bfseries in} $extractedWords$}
            \STATE $boundRect = word.contour.boundingRect$
            \IF{ $avgY > word.y$ {\bfseries and} $avgY < word.y + word.h$ }
            \STATE $line$.add $word$
            \STATE $noChange = false$
            \ENDIF
        \ENDFOR
        \UNTIL{$noChange = true$}
        \STATE $textlines$.add $newline$
   \UNTIL{$words$.empty $= true$}
\end{algorithmic}
\end{algorithm}

This algorithm fares well processing text images, where the text consists of parrallel text lines.
Angeled parallel text lines can be compensated for by rotating and padding the thresholded text image with different angles, and comparing the results.

Combinations of parrallel and slanted text lines however are not handled well by the algorithm, and will require further adaptations or more advanced algorithms to map the words to text lines correctly.

>>>>

\subsection{Contouring}
Our text segmentation approach is based on finding contours in the text image.
We can find these contours using %TODO:: reference paper contouring suzuki85
Through postprocessing we can recover the original words from these extracted contours.

\subsection{Postprocessing}
We can remove the incorrect contours generated by the contouring algorithm through postprocessing.
Secondly we will reconnect words that were split because they are not written in a single stroke.

\subsubsection{Contour removal}
The contouring algorthm can create incorrect contours because of left-over noise in the image.
The tool we will use to remove these incorrect contours is the bounding box we can draw around a contour.
This bounding box is the smallest rectangle containing said contour, and it will allow us to easily compare contour positions in the text image.

Firstly we remove the contours of which the bounding box is contained by the bounding box of a second contour.
We can do this because the text segmentation process will return the word in the image contained by its bounding box, so we will not change  % TODO:: NOTATION

Secondly we remove contours that are too small to contain parts of a word.
The contouring algorithm will generate contours of noise and punctuation.
These contours we do not want to process, so we will filter them out.

This is done by taking the average height of all the found contours in the text.
Iteratively we remove remove the contours with a height lesser than the a predetermined threshold,
\vspace{-5px}
\begin{equation}
    threshold = averageHeight * removalMultiplier
\end{equation}
with the $removalMultiplier$ initialized on $2 / 3$ as chosen value.

\subsubsection{Contour combination}
With the excess contours removed, we begin to reconstruct the original words from the text.
We do this by combining multiple contours into one where a word was separated into multiple parts.
We have created a straightforward overlapping algorithm, that combines word segments based on the overlap between the bounding box of their contours.

We have a parameter for the allowed gap between parts of the same word, and one for the minimal y-axis overlap of both their boundingboxes.

\begin{equation} %TODO:: FIX INDENTATION
    if \\ 
        wordXaxisGap < \\
            averageWordWidth * (1 / (averageCharacterCount * 2) )\\
        and \\
        wordYaxisOverlap > \\
        min(word1.height, word2.height) * requiredYaxisOverlap\\
    then \\
        combineWords
\end{equation}
The optimal parameters we found are:

\noindent 
$averageCharacterCount = 8$ and

\noindent
$requiredYaxisOverlap = 1/2$

With these conditions fulfilled, we combine both contours into one single contour, containing the found word.

\begin{algorithm}[tb]
   \caption{Algorithm for line reconstruction}
   \label{alg:linereconstr}
\begin{algorithmic}
   \STATE {\bfseries Input:} $extractedWords$, $textlines$
   \REPEAT
        \STATE $highestWord = findHighestWord $
        \STATE $newline = [highestWord]$
        \STATE Initialize $noChange = true$.
        \STATE $avgY= newline$.findAverageYpos
        \REPEAT
            \FOR{$word$ {\bfseries in} $extractedWords$}
            \STATE $boundRect = word.contour.boundingRect$
            \IF{ $avgY > word.y$ {\bfseries and} $avgY < word.y + word.h$ }
            \STATE $line$.add $word$
            \STATE $noChange = false$
            \ENDIF
        \ENDFOR
        \UNTIL{$noChange = true$}
        \STATE $textlines$.add $newline$
   \UNTIL{$words$.empty $= true$}
\end{algorithmic}
\end{algorithm}

\subsection{Line selection}
With the words extracted from the text, the next goal is to reconstruct the original text.
We will need to order these words into lines 
For the line selection process, an iterative algorithm, Algorithm \ref{alg:linereconstr}, is used.

@@
